---
title: "Tidy data and statistical visualisation <br> to support exploration of temporal data"
type: "Workshop"
author: "Earo Wang"
date: "May 29, 2017 <br> <br> Slides on http://bit.ly/tsvis17"
output:
  xaringan::moon_reader:
    css: ["default", "myremark.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r initial, echo = FALSE, cache = FALSE, results = 'hide'}
library(knitr)
options(htmltools.dir.version = FALSE, tibble.width = 60)
opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
  fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center', 
  fig.width = 12, fig.height = 9, fig.show = 'hold', 
  cache = TRUE, external = TRUE, dev = 'CairoPNG', dpi = 300
)
read_chunk('src/code.R')
```

## Bio

---

## Agenda

* Tidy temporal data
    + `ts` class vs `tibble` class
* What does `gg` in `ggplot2` stand for? A grammar of graphics
* How do we know if what we see really is there? Visual statistical inference
* Calendar-based view for visualising people's daily schedules
* Interactive data visualisation on the web
    + `plotly::ggplotly()`
    + Shiny

```{r load-library}
```

```{r theme-remark}
```
---

class: inverse middle center

# Tidy temporal data

---

.left-column[
## Melbourne pedestrian data .red[*]
### - sensor data

.footnote[.red[*] sourced from [the city of Melbourne](http://www.pedestrian.melbourne.vic.gov.au)]
]
.right-column[

```{r ped-loc, echo = TRUE}
```
]

---

.left-column[
## Melbourne pedestrian data .red[*]
### - sensor info
### - sensor map
]
.right-column[

```{r ped-map}
```
]

---

.left-column[
## Melbourne pedestrian data .red[*]
### - sensor info
### - sensor map
### - hourly counts
]
.right-column[

```{r ped-wide, echo = TRUE}
```
]

---

.left-column[
## Melbourne pedestrian data .red[*]
### - sensor info
### - sensor map
### - hourly counts
### - `autoplot`
]
.right-column[

```{r ped-ts, echo = TRUE, fig.height = 10}
```
]

---

.left-column[
## Tidy data
### - `ts` class
]
.right-column[
example here
]

---
